An AI is Haunting Campus

### 2024-06-29 ai-haunting-campus/ai-haunting-campus.qmd
Fixed a troublesome footnote and created a word document to count words.

A key reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they have been presented quite deliberately as _everything tools_^[Almost as flexible as, well, a naturally intelligent human, but faster and supposedly cheaper.]. While hammers are understood to have a limited range of uses that mostly have to do with concentrating force, and word processors are generally only used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

This is an utterly unrealistic expectation (also see automated grading below). Ideally, students should have the opportunity to discuss their drafted work with instructors before final submission (or again, could ask members of their study groups to proof-read their work, among many other ways to get feedback before submission). A conscientious instructor will invest time reading and understanding the draft work, so as to be able to provide quality feedback. But there is a more insidious aspect about getting feedback from an AI...who, actually, provides the feedback in this case? Whose data are used to train the machine that provides the feedback? Who fine-tunes the machine that provides the feedback? Whose perspectives and views are used to provide the feedback, and can they ever be accountable for the feedback provided^[[Automation bias](https://en.wikipedia.org/wiki/Automation_bias), the human inclination to favor suggestions made by automated decision-making systems, means that even when a human takes responsibility for decisions, AI companies remain unaccountable for their role in influencing outcomes^[GenAI is not a toy, as Air Canada discovered when the company was found liable for "negligent misrepresentation", after a court denied claims by the airline that the chatbot it deployed was somehow responsible for its own actions as an effectively separate legal entity [@garcia_what_2024].]. A poignant example involving AI-based face recognition and online proctoring is related by @hill_accused_2022.]?

>>>>

A key reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they have been presented quite deliberately as _everything tools_^[Almost as flexible as, well, a naturally intelligent human, but faster and supposedly cheaper. [Here](https://www.youtube.com/watch?v=yUoj9B8OpR8) is OpenAi's Mira Murati talking about "AI Everywhere: Transforming Our World, Empowering Humanity"--just maybe not all of humanity: "some creative jobs maybe will go away, but maybe they shouldn't have been there in the first place". One can be sure she means the jobs that produced the materials that OpenAI stole to train its AIs.]. While hammers are understood to have a limited range of uses that mostly have to do with concentrating force, and word processors are generally only used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

This is an utterly unrealistic expectation (also see automated grading below). Ideally, students should have the opportunity to discuss their drafted work with instructors before final submission (or again, could ask members of their study groups to proof-read their work, among many other ways to get feedback before submission). A conscientious instructor will invest time reading and understanding the draft work, so as to be able to provide quality feedback. But there is a more insidious aspect about getting feedback from an AI...who, actually, provides the feedback in this case? Whose data are used to train the machine that provides the feedback? Who fine-tunes the machine that provides the feedback? Whose perspectives and views are used to provide the feedback, and can they ever be accountable for the feedback provided^[[Automation bias](https://en.wikipedia.org/wiki/Automation_bias), the human inclination to favor suggestions made by automated decision-making systems, means that even when a human takes responsibility for decisions, AI companies remain unaccountable for their role in influencing outcomes. GenAI is not a toy, as Air Canada discovered when the company was found liable for "negligent misrepresentation", after a court denied claims by the airline that the chatbot it deployed was somehow responsible for its own actions as an effectively separate legal entity [@garcia_what_2024].]? A poignant example involving AI-based face recognition and online proctoring is related by @hill_accused_2022.

### 2024-06-29 paper/paper.qmd
Fixed a troublesome footnote and created a word document to count words.

---
title: "An AI is haunting campus"
subtitle: "Some considerations about the use of AI for teaching and learning in post-secondary education"
author: Antonio Páez (McMaster University)
date: "`r Sys.Date()`"
format:
    pdf: 
      fontsize: 7pt
    html: 
      self-contained: true
      grid: 
        margin-width: 350px
execute: 
  echo: fenced
  freeze: auto
reference-location: margin
#citation-location: margin
bibliography:
  - "`r system('kpsewhich ../bibliography/bibliography.bib', intern=TRUE)`"
---

@@

::: {.column-margin}
Karl Marx wrote about a "spectre...haunting Europe--the spectre of communism"; GenAI looks awfully like the spectre of unregulated surveillance capitalism.
:::

Disclaimer: _All opinions expressed here are mine and any resemblance to those of my employer(s) is purely coincidental_. _No generative AI tool was used to research or write this essay, only good old-fashioned library and internet search skills_.

Note: _This essay was written as an open research project, and it can be found in the following repository:_ https://github.com/paezha/University-AI-Panel

License: [CC-BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)

![](by-sa.png){width=50}

@@

<!--
> Instead of working with cohorts of learners in grades or classes, individual learning agendas are developed based on an assessment of the skills, competencies and capabilities the learner already has as well as ongoing conversations about the purpose of the learning. Where does the student want the learning to take them? AI systems, coupled with coaching, mentoring and some instruction and peer support networks then ensure the learner is supported on the carefully mapped learning journey. Teaching can still take place: Students challenged with a set of related issues can be brought together for a boot camp or focused period of instruction before returning to their individualized study activities.

>An end to “time” as a learning metric. Currently, undergraduate courses are described in terms of credit hours. Learning and assessment on demand, fast-tracking or slow-tracking using adaptive assessment, and modular learning when the student is ready liberates learners from the time-based metric.

> An end to exams as we know them. Exams are an efficient way to judge the output of batch teaching but are not good indicators of learner knowledge, skills and capabilities. With AI-enabled assessment generation, automated marking and assessment on demand, exams are no longer needed. More authentic assessments that are project-based, team challenges, work-based learning activities, and student self-directed projects can become the norm.

>Increased feedback and advising. AI systems do not have “office hours.” They are available 24x7 and never tire of finding new ways of responding to student queries and needs.Subject focused chatbots which have been “fed” all necessary information about a subject will do an excellent job of subject-matter advising and teaching.  Analytics suggest that some students need to be supported by teachers or others with special skills, such as writing coaches, statistical support workers and numeracy advisors. The aim will be to ensure every learner follows and is successful in the learning journey they have contracted to complete.

> Greater accessibility. Higher education systems are organized around blocks of time: semesters or terms. There is no longer good reason for this. Students should be able to begin their learning at any time and be assessed when they are ready. A four-year undergraduate degree could easily be completed in 16 months by a diligent student, for example, or in even less time by someone arriving in the program with substantial life and work experience. AI systems can also leverage technologies to ensure the resources and supports for learners reflect who they are (culture, language, experience and skill) and what challenges they face (disabilities, learning styles, etc.).

## As an aside {#as-an-aside}

On antropomorphizing GenAI tools. This happens mostly in ways that are barely liminal. For example, in the document "Overview of AI at McMaster" we read that GenAI's somehow have an "ability to engage in nuanced conversation". This phrasing could be simply a literary device, but even so it is indicative of a tendency to ascribe human attributes to a statistical model. Doing this risks amplifying existing biases. But in other cases, the effort is more blatant. Wrapping a statistical model in a human-like skin may actually be more devious, as in the new, sexy ChatGPT interface:

<video width="320" height="240" controls>
<source src="video/open-ai-commercial.mp4" type="video/mp4">
</video>

**NOTES:** Need to research more the argument that the interface is so important in the case of these tools because what they do well is trivial and what they do poorly is what is being hyped. Found the article [@farrugia_known_2024]; see here: https://fasterandworse.com/known-purpose-and-trusted-potential/

Are GenAI reality-warping machines? Think about the small and large ways in which these tools can be used to rewrite history, for relatively banal commercial objectives (monetizing clicks) or worse, for malicious propaganda campaigns.

**NOTE** Russia copied millions of articles from wikipedia wholesale, to replace the open encyclopedia with a home-grown version without all the bits that the current regime finds objectionable [@corfield_russia_2023; @jankowicz_russia_2023].

>The first step in liquidating a people,' said Hubl, 'is to erase its memory. Destroy its books, its culture, its history. Then have somebody write new books, manufacture a new culture, invent a new history. Before long the nation will begin to forget what it is and what it was. The world around it will forget even faster.  
>-― Milan Kundera, The Book of Laughter and Forgetting

Potential for weaponization of GenAI. **Farfetched (?) idea** Could it be that GenAI will be at some point declared a military-grade technology with restricted applications? (see https://www.defenseone.com/technology/2024/05/ai-and-other-tech-creating-special-operations-renaissance/396388/)

Peter Thiel on "AI is a military technology" https://www.businessinsider.com/peter-thiel-claim-ai-primarily-military-technology-experts-disagree-2019-8

AI industry and the military: @haskins_im_2024
-->

## Appendices

### AI Committee announcement {#ai-committee-announced}

This is the announcement of the launch of the AI Committee (see email from the Office of the Provost, dated March 20, 2024):

<!--https://app.fastmail.com/mail/search:expert+panels/T2d9b9367655ae218.M6205d28844daa67dfd3432ff-->

> "McMaster launches AI committee
>
> Whether you are a researcher, faculty member, student, administrator, or have another role on campus, you are likely thinking about Artificial Intelligence (AI) and its challenges and potential.
>
> AI is a powerful tool and I believe that universities have an important role to play in its adoption. At McMaster, we are well placed to prioritize an ethical approach to shaping policies and developing new ways of using AI to ensure a responsible and beneficial integration where it makes sense.
>
> I’m pleased to announce that McMaster has launched an AI Advisory Committee to explore the use of AI across campus. Three subcommittees have also been started to examine the use of AI in teaching and learning, research and operational excellence in our work. It is a strategy that makes McMaster one of the few institutions taking a pan-university approach to AI.
>
> Our earlier work to develop provisional guidelines for the use of generative AI in teaching and learning drew interest from other post-secondary institutions, which adopted McMaster's guidelines after we made them available through a Creative Commons licence.
>
> The committee will be led by three co-chairs, each bringing different expertise. They are Matheus Grasselli, deputy provost, Gayleen Gray, associate vice-president and chief technology officer and Gianni Parise, acting deputy vice-president, Research. They will be supported by Erin Aspenlieder, special advisor to the provost on generative AI.
>
> The three subcommittees will each by led by a subject matter expert. Kim Dej, vice-provost, Teaching and Learning, will lead the exploration of AI in teaching and learning, Maggie Pooran, executive director, Health, Safety, Well-Being & Labour Relations, will explore its use in operational excellence, and Martin Horn, associate dean, Faculty of Humanities, will look at implications for research.
>
> Each expert subcommittee will have a pool of contributors made up of members of the McMaster community. Depending on the initiative, some or all members will be invited to complete the specific work based on their expertise, interest and availability.
>
> If you would like to be considered for the expert panels, please complete this expression of interest form."

### AI Teaching and Learning Panel priority initiatives {#ai-panel-tasks}

The [Terms of Reference](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/ai-advisory-committee/) of the Artificial Intelligence Advisory Committee identify the following priority initiatives for the Expert Panel on Teaching and Learning:

- Provide recommendations to the AI Advisory Committee on updates to the Provisional Guidelines on the Use of Generative AI in Teaching and Learning.  
- Review and recommend training, resources and professional development on AI literacy for both educators and students (e.g. grants, microcredentials, guidebooks).  
- Engage in dialogue with both students and educators to understand their perspectives on AI in learning.  
- Endorse ethical standards and guidelines for AI usage in teaching environments.  
- Review and provide insights into assessments or tools that leverage AI for student evaluation.  
- Engage in cross-institutional dialogues to share insights and learn from global best practices.  
- Advocate for student involvement in the design and critique of AI-enabled teaching tools.

>>>>

---
title: "An AI is Haunting Campus"
format: docx
bibliography:
  - "`r system('kpsewhich ../bibliography/bibliography.bib', intern=TRUE)`"
---

@@

@@
